### 2022-12-17 abstract.tex
paper 6

% The credibility of a scientific finding is based on the fact that it can be
% replicated when the same study is repeated with new subjects.
% An important aspect of the credibility of a scientific finding is that the
% finding can be replicated when the same study is repeated with new subjects.
% An important aspect of the credibility of a scientific finding is its
% replicability -- the ability to produce similar results when the same study is
% repeated with new subjects.
An important aspect of the credibility of a scientific finding is its
replicability -- the ability that a similar finding can be obtained when the
same study is repeated with new subjects. However, various failures to replicate
major scientific findings in the social and life sciences indicate that
replicability is often lower than expected. This ``replication crisis'' has led
to several methodological reforms in the past decade, an increased conduct of
replication studies being one of them. Yet despite this increase in replication
studies, there is no consensus on a fundamental question: How should replication
success be defined statistically?
% Various methods already exist and various new ones have been proposed in
% response to the crisis. This thesis centres around a proposal based on the
% reverse-Bayes approach. The key idea is to reverse the traditional ``forward''
% use of Bayes' theorem (``prior + likelihood $\rightarrow$ posterior''),
% starting instead with a pre-specified posterior and deducing the corresponding
% prior (``posterior + likelihood $\rightarrow$
% prior''). % As such, the reverse-Bayes approach naturally fits to the replication
% % setting as it enables the mathematical formalization of scepticism regarding the
% % finding from the original study.
% This allows us to challenge the original finding by determining a sceptical
% prior for the underlying effect size such that the resulting posterior is no
% longer convincing. This prior then represents the position of a sceptic who
% remains unconvinced by the original study. Whether or not the sceptics'
% position is justified can then be assessed in light of the new data from the
% replication study. The larger the conflict between the replication data and
% the sceptic, the larger the degree of replication success. This procedure can
% be summarized in a single quantitative measure of replication success, termed
% the sceptical $p$-value.

Possible answers to this question are given in the first and largest part of
this thesis. It consists of three extensions to a recently proposed
reverse-Bayes method for quantifying replication success. The key idea of the
method is to challenge the data from the original study with a sceptical prior
distribution so that the resulting posterior distribution is no longer
``significant''. The goal of the replication study is then to show that the
sceptical prior is unrealistic, and replication success is achieved if there is
sufficient conflict between the prior and the replication data. The procedure
can be summarized in a single quantitative measure, termed the sceptical
$p$-value.
% extension 1
The first extension recalibrates the sceptical $p$-value to produce more
appropriate inferences in terms of effect size. The recalibration is chosen such
that for original studies which were borderline significant, replication success
can only be achieved if the effect estimate from the replication study is larger
than the effect estimate from the original study. We find that the recalibrated
sceptical $p$-value has good frequentist properties comparable to the standard
method used in practice.
% extension 2
The second extension replaces tail probabilities with Bayes factors as measures
of evidence. The procedure can again be summarized in a single measure called
the sceptical Bayes factor.
% The sceptical prior is now determined such that the original finding is no
% longer convincing in terms of a Bayes factor. In contrast to tail
% probabilities,
% We find that the extended method combines two notions of replicability: It
% requires that original and replication study independently provide evidence for
% an effect and ensures that their effect estimates are compatible.
The sceptical Bayes factor has similar properties as the sceptical $p$-value but
it avoids a statistical paradox which the sceptical $p$-value suffers from -- in
some situations the sceptical $p$-value may indicate replication success even
though the effect estimate from the replication is much smaller than the effect
estimate from the original study. % , which cannot happen with the sceptical
% Bayes factor.
% Bayes factors have a more
% natural interpretation and allow for direct quantification of evidence for one
% hypothesis versus another. As with the sceptical $p$-value, the procedure leads
% to a single measure quantifying the degree of replication success called the
% sceptical Bayes factor.
The third extension provides a framework for Bayesian design of replication
studies which allows for combining data from the original study with external
knowledge. This approach can lead to potentially more efficient designs compared
to classical approaches, and it can be used for replication design based on both
the sceptical $p$-value and the sceptical Bayes factor.
% External knowledge, such as data from other studies, is ubiquitous in the
% replication setting and taking it into account can lead to potentially more
% efficient designs compared to classical approaches.

@@

The last part of this thesis takes a meta-scientific perspective on
methodological research. Questionable research practices, such as selective
reporting of results, are often seen as main cause for replicability issues in
the medical and social sciences. These practices can similarly harm
methodological research, but are often not recognized. To raise awareness, an
illustrative simulation study is conducted in which it is shown how a novel
method can easily be presented as superior over established competitor methods
if questionable research practices are employed. Finally, several
recommendations are given to alleviate these issues, most importantly the need
for simulation protocols.

>>>>

% The credibility of a scientific finding is based on the fact that it can be
% replicated when the same study is repeated with new subjects.
% An important aspect of the credibility of a scientific finding is that the
% finding can be replicated when the same study is repeated with new subjects.
% An important aspect of the credibility of a scientific finding is its
% replicability -- the ability to produce similar results when the same study is
% repeated with new subjects.
An important aspect of the credibility of scientific findings is their
replicability -- the ability that a similar finding can be obtained when the
same study is repeated with new subjects. However, various failures to replicate
major scientific findings in the social and life sciences indicate that
replicability is often lower than expected. This ``replication crisis'' has led
to several methodological reforms in the past decade, an increased conduct of
replication studies being one of them. Despite this increase in replication
studies, there is no consensus on a fundamental question: How should replication
success be defined statistically?
% Various methods already exist and various new ones have been proposed in
% response to the crisis. This thesis centres around a proposal based on the
% reverse-Bayes approach. The key idea is to reverse the traditional ``forward''
% use of Bayes' theorem (``prior + likelihood $\rightarrow$ posterior''),
% starting instead with a pre-specified posterior and deducing the corresponding
% prior (``posterior + likelihood $\rightarrow$
% prior''). % As such, the reverse-Bayes approach naturally fits to the replication
% % setting as it enables the mathematical formalization of scepticism regarding the
% % finding from the original study.
% This allows us to challenge the original finding by determining a sceptical
% prior for the underlying effect size such that the resulting posterior is no
% longer convincing. This prior then represents the position of a sceptic who
% remains unconvinced by the original study. Whether or not the sceptics'
% position is justified can then be assessed in light of the new data from the
% replication study. The larger the conflict between the replication data and
% the sceptic, the larger the degree of replication success. This procedure can
% be summarized in a single quantitative measure of replication success, termed
% the sceptical $p$-value.

Possible answers to this question are given in the first and largest part of
this thesis. It consists of three extensions to a recently proposed
reverse-Bayes method for quantifying replication success. The key idea of the
method is to challenge the data from the original study with a sceptical prior
distribution so that the resulting posterior distribution no longer indicates
evidence for an effect. The goal of the replication study is then to show that
the sceptical prior is unrealistic, and replication success is achieved if there
is sufficient conflict between the prior and the replication data. The procedure
can be summarized in a single quantitative measure, termed the sceptical
$p$-value.
% extension 1
The first extension recalibrates the sceptical $p$-value so that replication
success takes effect size more appropriately into account. Specifically, the
recalibration is chosen such that for original studies which were borderline
significant, replication success can only be achieved if the effect estimate
from the replication study is larger than the effect estimate from the original
study. We find that the recalibrated sceptical $p$-value has good frequentist
properties comparable to the standard method used in practice.
% extension 2
The second extension replaces tail probabilities with Bayes factors as measures
of evidence. The procedure can again be summarized in a single measure called
the sceptical Bayes factor.
% The sceptical prior is now determined such that the original finding is no
% longer convincing in terms of a Bayes factor. In contrast to tail
% probabilities,
% We find that the extended method combines two notions of replicability: It
% requires that original and replication study independently provide evidence for
% an effect and ensures that their effect estimates are compatible.
The sceptical Bayes factor has similar properties as the sceptical $p$-value but
it avoids a statistical paradox which the sceptical $p$-value suffers from -- in
some situations the sceptical $p$-value may indicate replication success even
though the effect estimate from the replication is much smaller than the effect
estimate from the original study. % , which cannot happen with the sceptical
% Bayes factor.
% Bayes factors have a more
% natural interpretation and allow for direct quantification of evidence for one
% hypothesis versus another. As with the sceptical $p$-value, the procedure leads
% to a single measure quantifying the degree of replication success called the
% sceptical Bayes factor.
The third extension provides a framework for Bayesian design of replication
studies which allows for combining data from the original study with external
knowledge. This approach can lead to potentially more efficient designs compared
to classical approaches, and it can be used for replication design based on both
the sceptical $p$-value and the sceptical Bayes factor.
% External knowledge, such as data from other studies, is ubiquitous in the
% replication setting and taking it into account can lead to potentially more
% efficient designs compared to classical approaches.

@@

The last part of this thesis takes a meta-scientific perspective on
methodological research. Questionable research practices, such as selective
reporting of results, are often seen as main cause for replicability issues in
the medical and social sciences. These practices can similarly harm
methodological research, but are often not recognized. To raise awareness, an
illustrative simulation study is conducted in which it is shown how a novel
method can easily be presented as superior over established competitor methods
if questionable research practices are employed. Finally, several
recommendations are given to alleviate these issues.

### 2022-12-17 intro.Rnw
paper 6

@@

\begin{wrapfigure}{r}{0.27\textwidth}
  \centering
  \includegraphics[height = 6.9cm]{images/xkcdCrop.png}
  \caption*{Replication studies as \mbox{illustrated} by Randall Munroe
    \mbox{(\url{https://xkcd.com/242})}.}
\end{wrapfigure}

% How can we know if a finding from a scientific study is really true? For
% example, how can we know if the protective effect of a vaccine found in a study
% is real? The answer to such questions is of high importance to scientists,
% decision makers, and the general population. However, because any study result
% comes with uncertainty we can never know for sure.
The ability to trust scientific findings is essential for making evidence-based
decisions. But how can we know whether a scientific finding is trustworthy? For
instance, how can we know whether the protective effect of a vaccine found in a
study is indeed existent, or whether it is just a spurious finding? One way to
answer this question is through the use of \emph{replication studies}.

A replication study (or, simply, replication) involves repeating the original
study as closely as possible but with new subjects. If the replication yields
similar results to the original study, this increases our confidence in the
original finding. However, if the replication study yields conflicting results,
this lowers our confidence in the original finding. As such, replication studies
are an essential part of the scientific method. They provide a means for
substantiating genuine research findings and refuting findings which occurred
merely by chance. For this reason, ``successful'' replication is often a
requirement for acceptance of newly proposed scientific theories (\eg{} a new
physical model). Yet, replication studies are not only important for
researchers, but also for decision makers and the general population, as the
implementation of policies based on scientific findings may also depend on
whether these findings can be replicated (\eg{} market approval of a drug). The
results from replication studies may thus also have real-world consequences, and
it is therefore important that the methodological aspects of replication
studies, such as their statistical design and analysis, are well understood.

Despite their importance, the traditional academic system has made it
unattractive for researchers to conduct replication studies; the currencies of
science -- publications, citations, and grant money -- are typically easier to
acquire by conducting novel research. This is because until recently, most
scientific journals hardly published any replication studies as the emphasis was
on ``innovation'' rather than on ``repetition'' \citep{Makel2012, Martin2017,
  Coiera2021}. As a result, researchers often had no other choice than to focus
their efforts on producing novel and eye-catching research results in order to
build successful careers \citep{Binswanger2014, Moher2018}.

@@

\begin{table}[!htb]
  \centering
  \caption{Study-level summary statistics for an original and a replication
    study. The cumulative distribution function of the standard normal
    distribution is denoted by $\Phi(\cdot)$, and the $1 - \alpha$ quantile of
    the standard normal distribution is denoted by
    $\Phi^{-1}(1 - \alpha) = \z{\alpha}$. Confidence intervals and
    \textit{p}-values are based on the assumption that the effect estimate
    $\that$ is normally distributed around the unknown effect size $\theta$ with
    (known) variance equal to the squared standard error $\sigma^{2}$.}
  \label{tab:meta}
  \begin{tabular}{lcc}
    \toprule
    & Original study & Replication study \\
    \midrule
    effect estimate & $\that_o$ & $\that_{r}$ \\
    standard error & $\sigma_{o}$ & $\sigma_{r}$ \\
    $(1 - \alpha)$ confidence interval &
    % $[\that_{o} \pm \z{\alpha/2} \sigma_{o}]$ &
    % $[\that_{r} \pm \z{\alpha/2} \sigma_{r}]$ \\
    $[\that_{o} - \z{\alpha/2} \sigma_{o}, \that_{o} + \z{\alpha/2} \sigma_{o}]$ &
    $[\that_{r} - \z{\alpha/2} \sigma_{r}, \that_{r} + \z{\alpha/2} \sigma_{r}]$ \\
    $z$-value & $z_{o} = \that_{o}/\sigma_{o}$ & $z_{r} = \that_{r}/\sigma_{r}$ \\
    $p$-value (two-sided) & $p_{o} = 2\{1 - \Phi(|z_{o}|)\}$ & $p_{r} = 2\{1 - \Phi(|z_{r}|)\}$ \\
    % One-sided $p$-value & $p_{o} = 1 - \Phi(|z_{o}|)$ & $p_{r} = 2\{1 - \Phi(|z_{r}|)\}$ \\
    \bottomrule
  \end{tabular}
\end{table}

Most methods for analyzing replication studies can be formulated in terms of
study-level summary statistics as shown in Table~\ref{tab:meta}. % Effect
% estimates, standard error, confidence intervals, and $p$-values
All of these are routinely reported in research articles, and if one of them is
missing they can typically be calculated from the others. Using summary
statistics is also often the only possible way for conducting the replication
success analysis as the raw data from the original study may not be available to
the researchers conducting the replication study. The most important statistic
is the effect estimate $\that$. It provides an estimate of the underlying effect
size $\theta$ which quantifies the true effect or association of an
intervention/exposure with an outcome variable. Typical effect sizes are mean
differences and correlations (for continuous outcomes), odds ratios (for binary
outcomes), or hazard ratios (for time-to-event outcomes). Depending on the type
of effect size, a transformation might be required so that the assumption of
approximately normally distributed effect estimates around the unknown effect
size $\theta$ (for large enough sample sizes) is justified. This could be, for
instance, the Fisher $z$-transformation for correlations or the
log-transformation for odds/hazard ratios \citep[chapter 11]{Cooper2019}. The
associated standard error $\sigma$ represents the statistical uncertainty of the
estimate. Typically, the available standard error $\sigma$ is only an estimate
of the true standard error. However, in most parts of this thesis it will be
assumed that $\sigma$ corresponds to the true standard error. This is the same
assumption as in ordinary fixed-effect/random-effects meta-analysis, and
typically reasonable if the sample size of the studies is not too small. Under
the assumption of (asymptotic) normality, confidence intervals for $\theta$ and
$p$-values for testing the null hypothesis $H_{0} \colon \theta = 0$ can be
computed as shown in Table~\ref{tab:meta}.

>>>>

\begin{wrapfigure}{r}{0.29\textwidth}
  \centering
  \includegraphics[width = 0.2\textwidth]{images/xkcdCrop.png}
  \caption*{Replication studies as illustrated by Randall Munroe
    (\url{https://xkcd.com/242}).}
\end{wrapfigure}

@@

% How can we know if a finding from a scientific study is really true? For
% example, how can we know if the protective effect of a vaccine found in a study
% is real? The answer to such questions is of high importance to scientists,
% decision makers, and the general population. However, because any study result
% comes with uncertainty we can never know for sure.
Being able to trust scientific findings is essential for making evidence-based
decisions. But how can we know whether a scientific finding is trustworthy? For
instance, how can we know whether the protective effect of a vaccine found in a
study is indeed existent, or whether it is just a spurious finding? One way to
answer this question is through the use of \emph{replication studies}.

A replication study (or, simply, replication) involves repeating the original
study as closely as possible but with new subjects. If the replication yields
similar results to the original study, this increases our confidence in the
original finding. However, if the replication study yields conflicting results,
this lowers our confidence in the original finding. %  As such, replication studies
% are an essential part of the scientific method. They provide a means for
% substantiating genuine research findings and refuting findings which occurred
% merely by chance.
The concept of replicability has long been a central part of the scientific
method. For instance, \citet[p. 13]{Fisher1935} noted ``\textit{no isolated
  experiment, however significant in itself, can suffice for the experimental
  demonstration of any natural phenomenon}''. For this reason, ``successful''
replication is typically a requirement for acceptance of newly proposed
scientific theories (\eg{} a new physical model). Yet, replication studies are
not only important for researchers, but also for decision makers and the general
population, as the implementation of policies based on scientific findings may
also depend on whether these findings can be replicated. For instance, the
United States food and drug administration requires ``\textit{at least two
  adequate and well-controlled studies, each convincing on its own, to establish
  effectiveness}'' \citep[p. 3]{FDA1998} for market approval of a drug. The
results from replication studies may hence have real-world consequences. It is
therefore important that the methodological aspects of replication studies, such
as their statistical design and analysis, are well understood.

Despite their importance, the traditional academic system has made it
unattractive for researchers to conduct replication studies; publications,
citations, and grant money are typically easier to acquire by conducting novel
research. This is because until recently, most scientific journals hardly
published any replication studies as the emphasis was on ``innovation'' rather
than on ``repetition'' \citep{Makel2012, Martin2017, Coiera2021}. As a result,
researchers often had no other choice than to focus their efforts on producing
novel and eye-catching research results in order to build successful careers
\citep{Binswanger2014, Moher2018}.

@@

\begin{table}[!htb]
  \centering
  \caption{Study-level summary statistics for an original and a replication
    study. The cumulative distribution function of the standard normal
    distribution is denoted by $\Phi(\cdot)$, and the $1 - \alpha$ quantile of
    the standard normal distribution is denoted by
    $\Phi^{-1}(1 - \alpha) = \z{\alpha}$. Confidence intervals and
    \textit{p}-values for $H_{0}: \theta = 0$ are based on the assumption that
    an effect estimate $\that_{k}$ is normally distributed around the unknown
    effect size $\theta$ with (known) variance equal to the squared standard
    error $\sigma^{2}_{k}$ for $k \in \{o, r\}$.}
  \label{tab:meta}
  \begin{tabular}{lcc}
    \toprule
    & Original study & Replication study \\
    \midrule
    effect estimate & $\that_o$ & $\that_{r}$ \\
    standard error & $\sigma_{o}$ & $\sigma_{r}$ \\
    $(1 - \alpha)\times100\%$ confidence interval &
    % $[\that_{o} \pm \z{\alpha/2} \sigma_{o}]$ &
    % $[\that_{r} \pm \z{\alpha/2} \sigma_{r}]$ \\
    $[\that_{o} - \z{\alpha/2} \sigma_{o}, \, \that_{o} + \z{\alpha/2} \sigma_{o}]$ &
    $[\that_{r} - \z{\alpha/2} \sigma_{r}, \, \that_{r} + \z{\alpha/2} \sigma_{r}]$ \\
    $z$-value & $z_{o} = \that_{o}/\sigma_{o}$ & $z_{r} = \that_{r}/\sigma_{r}$ \\
    $p$-value (two-sided) & $p_{o} = 2\{1 - \Phi(|z_{o}|)\}$ & $p_{r} = 2\{1 - \Phi(|z_{r}|)\}$ \\
    % One-sided $p$-value & $p_{o} = 1 - \Phi(|z_{o}|)$ & $p_{r} = 2\{1 - \Phi(|z_{r}|)\}$ \\
    \bottomrule
  \end{tabular}
\end{table}

Most methods for analyzing replication studies can be formulated in terms of
study-level summary statistics as shown in Table~\ref{tab:meta}. % Effect
% estimates, standard error, confidence intervals, and $p$-values
All of these are routinely reported in research articles, and if one of them is
missing they can typically be calculated from the others. Using summary
statistics is also often the only possible way for conducting the replication
success analysis as the raw data from the original study may not be available to
the researchers conducting the replication study. The most important statistics
are the effect estimates $\that_{o}$ and $\that_{r}$. They provide an estimate
of the underlying effect size $\theta$ which quantifies the true effect or
association of an intervention/exposure with an outcome variable. Typical effect
sizes are mean differences and correlations (for continuous outcomes), odds
ratios (for binary outcomes), or hazard ratios (for time-to-event outcomes).
Depending on the type of effect size, a transformation might be required so that
the assumption of approximately normally distributed effect estimates around the
unknown effect size $\theta$ (for large enough sample sizes) is justified. This
could be, for instance, the Fisher $z$-transformation for correlations or the
log-transformation for odds/hazard ratios \citep[chapter 11]{Cooper2019}. The
associated standard errors $\sigma_{o}$ and $\sigma_{r}$ represent the
statistical uncertainty of the estimates. Typically, the available standard
errors are only estimates of the true standard errors. However, in most parts of
this thesis it will be assumed that $\sigma_{o}$ and $\sigma_{r}$ correspond to
the true standard errors. This is the same assumption as in ordinary
fixed-effect/random-effects meta-analysis, and typically reasonable if the
sample size of the studies is not too small. Under the assumption of
(asymptotic) normality, confidence intervals for $\theta$ and $p$-values for
testing the null hypothesis $H_{0} \colon \theta = 0$ can be computed as shown
in Table~\ref{tab:meta}.
